\pagestyle{plain}

\vspace*{2cm}
\begin{center}
% \selectlanguage{portuguese}
% \Large \bf Agradecimentos
\selectlanguage{english}
\Large \bf Acknowledgments
\end{center}
\vspace*{1cm} \setlength{\baselineskip}{0.6cm}

First of all, I would like to thank both of my advisors, Professor Bernardo Ferreira and Professor Alysson Bessani, for not only allowing me to work on this project but also for their patience, guidance, knowledge, and support throughout the whole process of this thesis.

I would like to thank my friends and colleagues who accompanied me on this journey and made it less lonely and more enjoyable. Furthermore, I also want to express a special thanks to Robin Vassantlal for helping me with COBRA and other problems I had along the way, in which he was always very helpful. 

Most importantly, I want to thank my family for their unconditional support.

At last, I gratefully acknowledge the financial support from Fundação para a Ciência e Tecnologia (FCT) through the project SMaRtChain (ref. \href{https://doi.org/10.54499/2022.08431.PTDC}{2022.08431.PTDC}) and the LASIGE Research Unit (ref. \href{https://doi.org/10.54499/UIDB/00408/2020}{UIDB/00408/2020} and ref. \href{https://doi.org/10.54499/UIDP/00408/2020}{UIDP/00408/2020}).

To all, each one with their unique importance, thank you very much.

\LIMPA
\LIMPA

~
\vfill

% \selectlanguage{portuguese}
\selectlanguage{english}
\begin{flushright}\textit{To those who have guided and inspired me along the way.}\end{flushright}

\LIMPA
